Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-04-01 W13report.tex
Final changes from Andy including clearing up derivation from Russell

Similarly, the predicted uncertainty of the flux measurement is:
%
\[\sigma = \frac{\sqrt{\sum_{x,y}\sigma{^2(x,y)}\phi^2(x,y)}}{\sum_{x',y'}\phi^2(x',y')}\]
%
where $\sigma{^2(x,y)}$ is the is the per-–pixel variance of the
unfiltered image. This reduces to the following:
%{XXX is this correct? - does it agree with Lupton?}
\[\sigma = \frac{\sqrt{\sigma peak{_{filt}^2}}}{\sum_{x',y'}\phi^2(x',y')}\]
%
where $\sigma peak{_{filt}^2}$ is the value of the variance plane,
after filtering, at the peak of the source.

The center of each source is determined using an LSST adaptation of
the SDSS centroiding algorithm "centroid.sdss". In general the source
will not be centered on a pixel, so "flux.filtered" uses a Lanczos 4
resampling kernel to shift the source to be centered on the nearest
pixel before applying the equations above.

@@

The main production runs included: an outer loop over the use of
pre--filtering or post--filtering for detection; variations of the
complexity of the Kernel basis, including the number of Gaussians
(nGauss = 1,3) and the degree of the modifying Laguerre polynomials
(degGauss = 1..6); and the spatial order of the Chebyshev polynomials
on $a_i(x,y)$ (spatialKernelOrder = 1..6).  The final production run
was labeled {\bf production10}, and data are available on the NCSA compute
cluster\footnote{/nfs/lsst8/becker/sparse\_diffim/production10\_sparse}.

{XXX it is hard to work through all of these permutations in the text
  and keep them in your head. Why not a table of the permutations}

@@

After running an ensemble of 144 pipeline configurations, we
aggregated the numbers of false detections.  We report here the results for
each of the 3 visits, and for pre-- and postfiltering for
detection.  We cut on {\tt flags.pixel.edge} to remove spurious
detections on the immediate border of the image; other than that we do
no other cuts other than requiring that the {\tt centroid.sdss} is
finite.  We separated the {\tt DiaSources} that match with the reference
catalog (3'' matching radius) from those that don't (referred to as
orphans) to investigate the origins of the detections.  We found no
significant correlation of false detections rate with sensor within the
raft (although the per--sensor optimal configs may be different, Section~\ref{sec-stats}); 
therefore we report the aggregate results for all 9 sensors in
the analyses that follow.

@@

\section{Results: Correlation with KernelCandidate Statistics \label{sec-stats}}

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false detections than positive false detections.  This drove an investigation 
into the symmetry of detection and measurement on positive/negative sources.  To do this we looked at the 
results of the image difference task and compared it to the same runs when the difference image was inverted
(multiplied by -1) before measurement.  The results are as follows:
\begin{itemize}
\item Centroids of negative sources always have integer or half--integer values.
\item The {\tt DiaSource} lists are always the same length.  {XXX dont
    understand this}
\item Centroiding on negative sources fails almost all the time: 99.7\% sources with negative flux
have the {\tt badcentroid} flag set.
\item psfFlux can be measured at 0. even when no flags are set.  {XXX dont
    understand this}
\item psfFlux can be measured as NaN with valid centroid values and no other flags set.
\item Fitted psfFlux can have the same sign in both the initial and inverted image in 0.1\% of measurements.
\item Some sources have the {\tt EDGE} bit set at one polarity but not the other.  
  This leads to a difference in the reported number of false positives as we only consider 
  those without the {\tt EDGE} flag set.
\end{itemize}

>>>>

If we define

\[w(x,y) \equiv \frac{\phi(x,y)}{\sum_{(x',y')}\phi^2(x',y')}\]

To calculate the variance, we assume a point source model for the intensity

\[I(x,y) = A\phi(x,y) + \epsilon(x,y)\]

\[\langle flux \rangle = \langle \sum_{(x,y)} w(x,y)(A\phi(x,y) +\epsilon(x,y))\rangle\]

Since $\langle \epsilon(x,y) \rangle = 0$ and $\langle \epsilon(x,y) * \epsilon(x',y') \rangle = \sigma^2(x,y)\delta$,

\[\langle flux \rangle = \sum_{(x,y)} w(x,y)A\phi(x,y)\]
and
\[Var(flux) \equiv \langle(flux - \langle flux \rangle)^2\rangle\]
\[Var(flux) = \left\langle{\left(\sum_{(x,y)} w(x,y)\epsilon(x,y)\right)^2}\right\rangle\]
\[Var(flux) = \sum_{(x,y)} w^2(x,y)\sigma^2(x,y)\]
Substituting $w(x,y)$
\[Var(flux) = \frac{\sum_{(x,y)}\phi^2(x,y)\sigma^2(x,y)}{(\sum_{(x',y')} \phi^2(x,y))^2}\]
This implies that the error $\sigma_{filt}$ on the flux at the location of $Ipeak_{filt}$ is
\[\sigma_{filt} = \frac{\sqrt{(\sum_{(x,y)} \phi^2(x,y)\sigma^2(x,y))}}{\sum_{(x',y')} \phi^2(x,y)}\]
%Similarly, the predicted uncertainty of the flux measurement is:
%
%\[\sigma = \frac{\sqrt{\sum_{x,y}\sigma{^2(x,y)}\phi^2(x,y)}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma{^2(x,y)}$ is the is the per--pixel variance of the
%unfiltered image. This reduces to the following:
%{XXX is this correct? - does it agree with Lupton?}
%\[\sigma = \frac{\sqrt{\sigma peak{_{filt}^2}}}{\sum_{x',y'}\phi^2(x',y')}\]
%
%where $\sigma peak{_{filt}^2}$ is the value of the variance plane,
%after filtering, at the peak of the source.

In practice, the numerator evaluates to the square root of the value in the variance plane
at the location of the peak flux.

The center of each source is determined using an LSST adaptation of
the SDSS centroiding algorithm ``centroid.sdss". In general the source
will not be centered on a pixel, so ```flux.filtered" uses a Lanczos 4
resampling kernel to shift the source to be centered on the nearest
pixel before applying the equations above.

@@

The main production runs included: an outer loop over the use of
pre--filtering or post--filtering for detection; variations of the
complexity of the Kernel basis, including the number of Gaussians
(nGauss = 1,3) and the degree of the modifying Laguerre polynomials
(degGauss = 1..6); and the spatial order of the Chebyshev polynomials
on $a_i(x,y)$ (spatialKernelOrder = 1..6).  This results in $2*2*6*6=144$ permutations
of the configs.  See the following pseudocode
for how the loops iterate over the configs.

\clearpage
\begin{verbatim}
for prefilter in (True, False):
    for nGauss in (1,3):
        for degGauss in xrange(1,7):
            for order in xrange(1,7):
                runDiffimPipelin(prefilter, nGauss, degGauss, order)
\end{verbatim}

The final production run
was labeled {\bf production10}, and data are available on the NCSA compute
cluster\footnote{/nfs/lsst8/becker/sparse\_diffim/production10\_sparse}.

@@

After running an ensemble of 144 pipeline configurations, we
aggregated the numbers of false detections.  We report here the results for
each of the 3 visits, and for pre-- and postfiltering for
detection.  We cut on {\tt flags.pixel.edge} to remove spurious
detections on the immediate border of the image; other than that we do
no other cuts other than requiring that the {\tt centroid.sdss} is
finite.  We separated the {\tt DiaSources} that match with the reference
catalog (3'' matching radius) from those that don't (referred to as
orphans) to investigate the origins of the detections.  We found no
significant correlation of false detections rate with sensor within the
raft (although the per--sensor optimal configs may be different, Section~\ref{sec-resstats}); 
therefore we report the aggregate results for all 9 sensors in
the analyses that follow.

@@

\section{Results: Correlation with KernelCandidate Statistics \label{sec-resstats}}

@@

\subsection{Detection and Measurement on Negative Sources}
We see more negative false detections than positive false detections.  This drove an investigation 
into the symmetry of detection and measurement on positive/negative sources.  To do this we looked at the 
results of the image difference task and compared it to the same runs when the difference image was inverted
(multiplied by -1) before measurement.  The results are as follows:
\begin{itemize}
\item Centroids of negative sources always have integer values.
\item Whether detected on the original difference image or the inverted image, the {\tt DiaSource} lists are the same length in every
case.  That is to say {\tt len(afwTable.SourceCatalog.readFits(\it{path}))} returns the same value in both cases.  We take this to mean
that the same objects trigger detection in both situations although this cannot be directly varified since the objects with failed centroids are
not the same in both lists.  
\item Centroiding on negative sources fails almost all the time: 99.7\% sources with negative flux
have the {\tt badcentroid} flag set.
\item In some instances psfFlux is stored as 0.0 (not {\tt NaN}) without any centroid or photometry flag set in the {\tt Source} object.  This
happens more than we would expect if it was just a chance event.  This indicates an uncaught failure mode.
\item psfFlux can be measured as NaN with valid centroid values and no other flags set.
\item Fitted psfFlux can have the same sign in both the initial and inverted image in 0.1\% of measurements.
\item Some sources have the {\tt EDGE} bit set at one polarity but not the other.  
  This leads to a difference in the reported number of false positives as we only consider 
  those without the {\tt EDGE} flag set.
\end{itemize}
